---
title: "Power calculation for alteplase RDD"
author: "Patrick Lawlor"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(pwr)
```

Effect sizes reported in the literature (from UpToDate and original studies):

Between 3 and 4.5 hours:

* ECASS 3: Favorable outcome in 52% vs 45%.
* Pooled data/meta-analysis: Favorable outcome in 35% vs 30%

After 4.5 hours:

* Pooled data/meta-analysis: Favorable outcome in 33% vs 31%

Fuzzy effect in data: 

* Cohort A (pre-2008, 3 hour threshold): 66.5% got tPA before threshold vs 27.0% after threshold
* Cohort B (post-2009, 4.5 hour threshold): 36.3% got tPA before threshold vs 13.0% after threshold
* Cohort C (post-2009, 3 hour threshold): 87.4% got tPA before threshold vs 73.0% after threshold


```{r}
prop1 <- 0.524 # From ECASS 3
prop2 <- 0.452 # From ECASS 3

compliance <- 0.395 # compliance = 0.665 - 0.27, from Cohort A

effect_size_h <- ES.h(prop1, prop2) # Calculate Cohen's h, based on arcsine transformation
effect_size_actual <- effect_size_h*compliance

design_effect <- 4.0 # typical design effect from RDD

alpha <- 0.05 # standard value
power <- 0.8 # standard value

power_calc <- pwr.2p.test(h = effect_size_actual, sig.level = alpha, power = power)
n <- power_calc$n*design_effect

sprintf("%g patients required per group",n)
```

